---
title: "DM2 project - WarGames report plots"
author: "Maria Ines Silva"
date: "2 June 2018"
output: 
  html_document: 
    theme: cosmo
    toc: yes
---

<style>

table, td, th {
  border: none;
  padding-left: 1em;
  padding-right: 1em;
  min-width: 50%;
  margin-left: auto;
  margin-right: auto;
  margin-top: 1em;
  margin-bottom: 1em;
}

</style>

***

```{r setup, include=FALSE}

knitr::opts_chunk$set(warning=FALSE,
                      message=FALSE,
                      echo=FALSE,
                      fig.align="center",
                      fig.path="./Output/")

## Importing packadges
packages <- c("dplyr", "ggplot2", "tidyr", "xlsx", "lubridate", "stringr",
              "scales","kableExtra", "corrplot", "robustbase", "moments")
lapply(packages, library, character.only = TRUE)

# NOVA-IMS grey : #5C666C
# NOVA-IMS green : #BED62F

```


# Data exploration and pre-processing

```{r input dataset}

raw_data <- read.xlsx2(file = "./Input/Group_14_WarGames_campaign.xlsx",
                       sheetIndex = 1,
                       colIndex = c(1:26,29),
                       colClasses = c("character", "integer", rep("character",2),
                                      "numeric", rep("integer",2), "Date",
                                      "integer", rep("numeric",6), 
                                      rep("integer",5), rep("character",7)),
                       stringsAsFactors=FALSE) %>% 
          mutate(DepVar=factor(DepVar)) %>% 
          select(-Custid)

raw_data$DepVar <- factor(raw_data$DepVar,levels = rev(levels(raw_data$DepVar)))

campaign_specs <- read.xlsx2(file = "./Input/Group_14_WarGames_campaign.xlsx",
                   sheetIndex = 1,
                   colIndex = c(27:28),
                   endRow = 2,
                   colClasses = rep("numeric",2))

```



## Summary statistics

### Target variable

The target variable for this analysis is `DepVar`. It is a binary variable, identifying whether or not the customer accepted the campaign's offer. In this dataset, `r percent(nrow(filter(raw_data, DepVar==1))/nrow(raw_data))` of the customers accepted the offer and there are no missing values. Bellow is the frequency plot of this variable:

```{r DepVar overview, fig.height=1.5, fig.width=1.5}

raw_data %>%
          select(DepVar) %>% 
          ggplot(mapping = aes(x=DepVar)) +
          geom_bar(fill="#5C666C")+
          labs(y="Frequency")+
          theme_minimal(base_size = 9) +
          theme(panel.grid=element_blank(),
                axis.title.x =element_blank())
```



### Categorical predictors

```{r cat overview plot 1, fig.height=3, fig.width=4.5}
raw_data %>% 
          select(AcceptedCmp2:DepVar) %>% 
          gather(key=variable,value=level,-DepVar) %>% 
          ggplot(mapping = aes(x=level,fill=DepVar)) +
          geom_bar()+
          theme_minimal(base_size = 9) +
          theme(panel.grid=element_blank(),
                axis.title = element_blank())+
          scale_fill_manual(values = c("0"="#5C666C", 
                                        "1"="#BED62F"))+
          facet_wrap(~variable,ncol=3, scales = "free_x")
```


```{r cat overview plot 2, fig.height=3, fig.width=2.6}
raw_data %>% 
          select(Education, Marital_Status, DepVar) %>% 
          gather(key=variable,value=level,-DepVar) %>% 
          ggplot(mapping = aes(x=level,fill=DepVar)) +
          geom_bar()+
          theme_minimal(base_size = 9) +
          theme(panel.grid=element_blank(),
                axis.title = element_blank())+
          guides(fill=FALSE)+
          scale_fill_manual(values = c("0"="#5C666C", 
                                        "1"="#BED62F"))+
          facet_wrap(~variable,ncol=1, scales = "free_x")
```



### Numeric predictors

```{r num overview table}
temp <- raw_data %>% 
          select(Year_Birth, Income:NumWebVisitsMonth,-Dt_Customer, DepVar) %>% 
          gather(key=variable,value=value,-DepVar) %>% 
          group_by(variable) %>% 
          summarise(mean=round(mean(value,na.rm=TRUE),2),
                    coef_var=round(sd(value,na.rm=TRUE)/mean(value,na.rm=TRUE),2),
                    skewness=round(skewness(value,na.rm=TRUE),2),
                    min=round(min(value,na.rm=TRUE),1),
                    max=round(max(value,na.rm=TRUE),1)) %>% 
          ungroup()

write.xlsx(temp,"./Output/explore - num stats.xlsx")

knitr::kable(temp,
             format="html",
             align="c",
             col.names = c("Product","Mean","Coef. of variation",
                           "Skewness",  "Minimum","Maximum")) %>%
          kable_styling(bootstrap_options = "striped",
                        full_width = F)
```

For the Dt_Customer variables, it does not make sense to compute the mean or the standard deviation. Therefore, we computed:

* Mean- `r mean(raw_data$Dt_Customer)`
* Minimum - `r min(raw_data$Dt_Customer)`
* Maximum - `r max(raw_data$Dt_Customer)`
* Median - `r median(raw_data$Dt_Customer)`

And here are the histograms for the annex:

```{r num overview plot 1, fig.height=7, fig.width=8}
raw_data %>% 
          select(Year_Birth, Income:NumWebVisitsMonth, DepVar, -Teenhome, -Kidhome) %>% 
          gather(key=variable,value=level,-DepVar) %>% 
          ggplot(mapping = aes(x=level,fill=DepVar)) +
          geom_histogram()+
          theme_minimal(base_size = 9) +
          theme(panel.grid=element_blank(),
                axis.title = element_blank(),
                legend.position = "none")+
          scale_fill_manual(values = c("0"="#5C666C", 
                                        "1"="#BED62F"))+
          facet_wrap(~variable,ncol=3, scales = "free")
```


```{r num overview plot 2, fig.height=1.5, fig.width=5}
raw_data %>% 
          select(Teenhome, Kidhome, DepVar) %>% 
          gather(key=variable,value=level,-DepVar) %>% 
          ggplot(mapping = aes(x=level,fill=DepVar)) +
          geom_bar()+
          theme_minimal(base_size = 9) +
          theme(panel.grid=element_blank(),
                axis.title = element_blank())+
          scale_fill_manual(values = c("0"="#5C666C", 
                                        "1"="#BED62F"))+
          facet_wrap(~variable,ncol=2, scales = "free")
```


## Outliers

```{r outliers threshold - Tukey modified}
temp <- raw_data %>% 
        select(Year_Birth, Income:NumWebVisitsMonth,-Dt_Customer,-Kidhome,-Teenhome) %>% 
        gather(key=variable,value=value) %>% 
        group_by(variable) %>% 
        summarise(medcouple=round(mc(value, na.rm=TRUE),2),
                  q1=quantile(value,probs=0.25,na.rm=TRUE),
                  q3=quantile(value,probs=0.75,na.rm=TRUE),
                  left_threshold_1=round(q1-1.5*exp(-4*medcouple)*(q3-q1),2),
                  left_outliers_1=percent(sum(value<left_threshold_1,na.rm=TRUE)/2500),
                  right_threshold_1=round(q3+1.5*exp(3*medcouple)*(q3-q1),2),
                  right_outliers_1=percent(sum(value>right_threshold_1,na.rm=TRUE)/2500)
        ) %>% 
        ungroup()

write.xlsx(temp,"./Output/outliers - Tukey Modified.xlsx")

knitr::kable(temp,
             format="html",
             align="c",
             col.names = c("Variable", "Medcouple","Q1", "Q3", "Left threshold",
                           "% of outliers", "Right threshold","% of outliers")) %>%
        kable_styling(bootstrap_options = "striped",
                      full_width = F)
```

```{r outliers plot 1, fig.height=2, fig.width=4}
temp <- raw_data %>% 
          mutate(Mnt=MntMiniatures+MntCard_Games+MntMagazines+
                              MntScenario+MntPainting_Material,
                 Frq=NumCatalogPurchases+NumStorePurchases+NumWebPurchases)

temp %>% 
        ggplot(mapping = aes(x=Income,fill=DepVar)) +
        geom_histogram() +
        geom_vline(xintercept=147000)+
        theme_minimal(base_size = 9) +
        theme(axis.title.y = element_blank())+
        scale_fill_manual(values = c("0"="#5C666C", 
                                      "1"="#BED62F"))
```

```{r outliers plot 2, fig.height=2, fig.width=2.5}
temp <- raw_data %>% 
          mutate(Mnt=MntMiniatures+MntCard_Games+MntMagazines+
                              MntScenario+MntPainting_Material,
                 Frq=NumCatalogPurchases+NumStorePurchases+NumWebPurchases)

temp %>% 
        ggplot(mapping = aes(x=MntCard_Games, y=Mnt, color=DepVar)) +
        geom_point(alpha=0.8, size=1)+
        geom_vline(xintercept=1500)+
        theme_minimal(base_size = 9) +
        labs(y="Total amount spent")+
        theme(legend.position = "none")+
        scale_color_manual(values = c("0"="#5C666C", 
                                      "1"="#BED62F"))

temp %>% 
        ggplot(mapping = aes(x=NumDealsPurchases, y=Frq, color=DepVar)) +
        geom_point(alpha=0.8, size=1)+
        geom_vline(xintercept=14)+
        theme_minimal(base_size = 9) +
        labs(y="Total no. purchases")+
        theme(legend.position = "none")+
        scale_color_manual(values = c("0"="#5C666C", 
                                      "1"="#BED62F"))

temp %>% 
        ggplot(mapping = aes(x=NumCatalogPurchases, y=Frq, color=DepVar)) +
        geom_point(alpha=0.8, size=1)+
        geom_vline(xintercept=20)+
        theme_minimal(base_size = 9) +
        labs(y="Total no. purchases")+
        theme(legend.position = "none")+
        scale_color_manual(values = c("0"="#5C666C", 
                                      "1"="#BED62F"))

temp %>% 
        ggplot(mapping = aes(x=NumWebVisitsMonth, y=Frq, color=DepVar)) +
        geom_point(alpha=0.8, size=1)+
        geom_vline(xintercept=13)+
        theme_minimal(base_size = 9) +
        labs(y="Total no. purchases")+
        theme(legend.position = "none")+
        scale_color_manual(values = c("0"="#5C666C", 
                                      "1"="#BED62F"))
```


## Missing values

```{r missing values}

missings <- raw_data %>% 
        filter((is.na(Income) | is.na(MntBrandA._Material) | is.na(MntScenario)))
```

There are 3 variables with missing values - `Income`, `MntBrandA._Material` and `MntScenario`. There are `r nrow(missings)` customers with a missing values in at least on of these variables, which represents `r percent(nrow(missings)/2500)` of the raw dataset.


# Variable selection

```{r clean data import}
clean_data2 <- read.csv(file = "./Input/clean_data2_TRAIN.csv",stringsAsFactors=FALSE) %>% 
          mutate_at(vars(DepVar), funs(factor(.)))

colnames(clean_data2)[17:25] <- str_sub(colnames(clean_data2)[17:25],7)

clean_data2$DepVar <- factor(clean_data2$DepVar,levels = rev(levels(clean_data2$DepVar)))        
          
clean_data2 <- clean_data2 %>% 
          rename(MntBrandA_Material=IMP_MntBrandA__Material,
                 MntScenario=IMP_MntScenario,
                 Income=IMP_Income) %>% 
          select(-X_WARN_)
```


## Redundancy

```{r person corrplot, fig.height=2.5, fig.width=2.5}
clean_data2 %>% 
          select(Mnt, Frq, Income, average_purchase) %>% 
          cor(method="pearson") %>% 
          corrplot(method = "color",
                   type="upper",
                   addCoef.col = "black",
                   tl.col = "black",
                   tl.cex = 0.7,
                   cl.pos = "n",
                   number.cex = 0.7,
                   col=colorRampPalette(c("black", "black", "white","#5C666C"))(1000))
```

```{r spearman corrplot, fig.height=2.5, fig.width=2.5}
clean_data2 %>% 
          select(Mnt, Frq, Income, average_purchase) %>% 
          cor(method="spearman") %>% 
          corrplot(method = "color",
                   type="upper",
                   addCoef.col = "black",
                   tl.col = "black",
                   tl.cex = 0.7,
                   cl.pos = "n",
                   number.cex = 0.7,
                   col=colorRampPalette(c("black", "black", "white","#5C666C"))(1000))
```


The folllowing scatter plots shows a strange relation between `average_purchase` and `Frq`:

```{r relation frq-av_pur, fig.height=2.8, fig.width=3.5}
clean_data2  %>% 
          ggplot(mapping = aes(y=Frq, x=average_purchase, color=DepVar)) +
          geom_point(alpha=0.8, size=1)+
          theme_minimal(base_size = 9) +
          scale_color_manual(values = c("0"="#5C666C", 
                                        "1"="#BED62F"))
```


## Relevancy

```{r variable worth, fig.height=2.5, fig.width=6}
var_worth <- read.csv(file = "./Input/variable_worth.csv",stringsAsFactors=FALSE)

var_worth$Variable <- factor(var_worth$Variable,
                             levels = var_worth$Variable[order(var_worth$Worth, decreasing = TRUE)])

var_worth %>% 
        ggplot(mapping = aes(x=Variable,y=Worth)) +
        geom_bar(stat="identity",fill="#5C666C")+
        #coord_flip()+
        theme_minimal(base_size = 9) +
        theme(panel.grid=element_blank(),
              axis.title.x = element_blank(),
              axis.text.x = element_text(angle = 90, hjust = 1))
```


## Final Variables

```{r final vars 1, fig.height=3, fig.width=7.5}
clean_data2 %>% 
          select(average_purchase, Frq, Recency, MntCard_Games,
                 NumCatalogPurchases, NumStorePurchases,
                 NumDealsPurchases, NumWebPurchases, DepVar) %>% 
          gather(key=variable,value=level,-DepVar) %>% 
          ggplot(mapping = aes(x=level,fill=DepVar)) +
          geom_histogram()+
          theme_minimal(base_size = 9) +
          theme(panel.grid=element_blank(),
                axis.title = element_blank(),
                legend.position = "none")+
          scale_fill_manual(values = c("0"="#5C666C", 
                                        "1"="#BED62F"))+
          facet_wrap(~variable,ncol=4, scales = "free")
```

```{r final vars 2, fig.height=1.5, fig.width=6}
clean_data2 %>% 
        select(campaign_acceptance, Marital_Status, DepVar) %>% 
        gather(key=variable,value=level,-DepVar) %>% 
        ggplot(mapping = aes(x=level,fill=DepVar)) +
        geom_bar()+
        theme_minimal(base_size = 9) +
        theme(panel.grid=element_blank(),
              axis.title = element_blank())+
        scale_fill_manual(values = c("0"="#5C666C", 
                                     "1"="#BED62F"))+
        facet_wrap(~variable,ncol=3, scales = "free")
```


## Cross-validation results

```{r roc curve, fig.height=2.5, fig.width=5.5}
roc_stats <- read.csv(file = "./Input/Cross-validation results - roc curve.csv",
                      stringsAsFactors=FALSE) %>% 
        rename(Model=Model.Description)

roc_stats %>% 
        ggplot(mapping = aes(x=X1...Specificity,y=Sensitivity, color=Model)) +
        geom_line() +
        theme_minimal(base_size = 9)  +
        guides(color=guide_legend(ncol=2))+
        labs(x="1-Specificity",y="Sensitivity",color="Model:")+
          theme(panel.grid=element_blank())
```


```{r profit curve 1, fig.height=2.5, fig.width=3}

profit_curve <- read.xlsx2(file = "./Input/WarGame classification_ProfitCurves.xlsx",
                       sheetIndex = 1,
                       colClasses = c(rep("character",3), rep("numeric",9)),
                       stringsAsFactors=FALSE) %>% 
        transmute(model=MODELDESCRIPTION,
                  decile=DECILE/100,
                  profit=Profit,
                  response_rate=RESPC/100)


profit_curve %>% 
        ggplot(mapping = aes(x=decile, y=profit, color=model)) +
        geom_line()+
        geom_vline(xintercept=0.15,size=0.2)+
        labs(x="Customers' percentile",y="Profit",color="Model:")+
        scale_x_continuous(labels = scales::percent)+
        theme_minimal(base_size = 9) +
        guides(color=guide_legend(ncol=2))+
        theme(panel.grid=element_blank(),
              legend.position = "none")
```


```{r profit curve 2, fig.height=2.5, fig.width=6}
profit_curve %>% 
        ggplot(mapping = aes(x=decile, y=response_rate, color=model)) +
        geom_line()+
        geom_vline(xintercept=0.15,size=0.2)+
        labs(x="Customers' percentile",y="Response rate",color="Model:")+
        scale_x_continuous(labels = scales::percent)+
        scale_y_continuous(labels = scales::percent)+
        theme_minimal(base_size = 9) +
        guides(color=guide_legend(ncol=2))+
        theme(panel.grid=element_blank())
```
